### 2020-11-30 Manuscript/manuscriptPGPE.tex
Update manuscript

%% ****** Start of file aiptemplate.tex ****** %
%%
%%   This file is part of the files in the distribution of AIP substyles for REVTeX4.
%%   Version 4.1 of 9 October 2009.
%%
%
% This is a template for producing documents for use with 
% the REVTEX 4.1 document class and the AIP substyles.
% 
% Copy this file to another name and then work on that file.
% That way, you always have this original template file to use.

%\documentclass[aip,graphicx]{revtex4-1}
%\documentclass[aip,reprint]{revtex4-1}

%\usepackage{graphicx}

%\draft % marks overfull lines with a black rule on the right
%\documentclass[pre,aps,floatfix,authordate1-4,twocolumn]{revtex4-1}
%\documentclass[pre,aps,floatfix,authordate1-4]{revtex4-1}

\documentclass[aps,prl,superscriptaddress,twocolumn]{revtex4}

%\documentclass[aps,prl,preprint,groupedaddress]{revtex4}

\usepackage{rotating} 
\usepackage{times}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage[obeyFinal]{easy-todo}
\usepackage{csquotes}
\usepackage{xr}
\externaldocument{manuscriptPGPEsuppl}

\begin{document}

% Use the \preprint command to place your local institutional report number 
% on the title page in preprint mode.
% Multiple \preprint commands are allowed.
%\preprint{}

\title{NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PE and PG lipids} %Title of paper

% repeat the \author .. \affiliation  etc. as needed
% \email, \thanks, \homepage, \altaffiliation all apply to the current author.
% Explanatory text should go in the []'s, 
% actual e-mail address or url should go in the {}'s for \email and \homepage.
% Please use the appropriate macro for the type of information

% \affiliation command applies to all authors since the last \affiliation command. 
% The \affiliation command should follow the other information.

\author{Pavel Buslaev}
\affiliation{University of Jyv{\"a}skyl{\"a}}

\author{Fernando Favela-Rosales}
\affiliation{Departamento de Investigaci\'{o}n, Tecnol\'{o}gico Nacional de M\'{e}xico, Campus Zacatecas Occidente, M\'{e}xico}

\author{Patrick Fuchs}
\affiliation{Paris, France}

\author{Matti Javanainen}
\affiliation{Institute of Organic Chemistry and Biochemistry of the 
Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2, CZ-16610 Prague 6, Czech Republic}

\author{Jesper J. Madsen}
\affiliation{Department of Chemistry, The University of Chicago, Chicago, Illinois, United States of America}
\affiliation{Department of Global Health, College of Public Health, University of South Florida, Tampa, Florida, United States of America}

\author{Josef Melcr}
\affiliation{Institute of Organic Chemistry and Biochemistry of the 
Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2, CZ-16610 Prague 6, Czech Republic}
\affiliation{Groningen Biomolecular Sciences and Biotechnology Institute 
and The Zernike Institute for Advanced Materials, 
University of Groningen, 9747 AG Groningen, The Netherlands}

\author{Markus S. Miettinen}
% \affiliation[Max Planck Institute of Colloids and Interfaces]{Department of Theory and Bio-Systems, Max Planck Institute of Colloids and Interfaces, 14424 Potsdam, Germany}
\affiliation{Department of Theory and Bio-Systems, Max Planck Institute of Colloids and Interfaces, 14424 Potsdam, Germany}

\author{O. H. Samuli Ollila}
\email[]{samuli.ollila@helsinki.fi}
\affiliation{Institute of Biotechnology, University of Helsinki}

\author{Chris G. Papadopoulos}
\affiliation{I2BC - University Paris Sud}

\author{Antonio Pe{\'o}n}
\affiliation{Spain}

\author{Thomas J. Piggot}
\affiliation{Chemistry, University of Southampton, Highfield, Southampton SO17 1BJ, United Kingdom}

\author{Pierre Poulain}
\affiliation{Paris, France}

% Collaboration name, if desired (requires use of superscriptaddress option in \documentclass). 
% \noaffiliation is required (may also be used with the \author command).
%\collaboration{}
%\noaffiliation

\date{\today}

\begin{abstract}
% insert abstract here
Abstract
\end{abstract}

%\pacs{}% insert suggested PACS numbers in braces on next line

\maketitle %\maketitle must follow title, authors, abstract and \pacs

% Body of paper goes here. Use proper sectioning commands. 
% References should be done using the \cite, \ref, and \label commands

%\label{}
\section{Introduction}

PE and PG lipids are most common lipids in bacteria \cite{sohlenkamp16}.
Zwitterionic PE is the second most abundant glycerophospholipid in eukaryotic cells
and has been related to the diseases \cite{vance15,calzada16,patel17}.
Anionic PG lipids are less abundant, but is also proposed to be fundamental for terrestrial life \cite{furse17}.
PE and PG affect membrane protein functionality \cite{hariharan18} and bind to various proteins \cite{yeagle14}.
PE headgroup is also prone for negative membrane curvature and causes membrane fusion \cite{Chernomordik08,calzada16}.
Therefore, the PE and PG headgroup structures play probably essential roles in 
many biological processes.

Structural details of lipid headgroups are mainly studied using NMR experiments, which
suggest that the glycerol backbone structures are largely similar irrespectively of the headroup \cite{gally81}, 
glycerol backbone and headgroup structure and behaviour are similar in model membranes and in bacteria \cite{gally81,scherer87,seelig90},
and the headgroup structures are similar in PC, PE and PG lipids, while headgroup is more rigid in PS lipids \cite{wohlgemuth80,buldt81}. 
%Extensive discussion about structural details of PE, PG or PS headgroups do not exists (as far as I know), 
%In contrast to PC lipids (see \cite{botan15} and references therein).
Some attempts to resolve conformational ensembles from NMR for PC and PE lipids have been made,
but lesser extend for PG or PS lipids \cite{seelig77c,davis83,Semchyschyn04}.
Classical molecular dynamics simulations could potentially give such ensembles and therefore enable
the detailed studies of lipid headgroup behaviour in complex biomolecular systems, but current
force fields are not accurate enough to reproduce the correct conformational ensembles for PC and PS headgroups \cite{botan15,antila19}.
Several MD simulations of PE and PG lipids have been published especially in the context of modeling
inner membrane of Gram-negative
bacteria \cite{devries04,murzyn05,pedersen06,zhao07,gurtovenko08,zhao08,henin09,kukol09,tsai12,dickson12,venable13,dickson14,berglund15}
\todo{There may be some relevant publication missing from here},
but evaluation of glycerol backbone and headgroup structures against experiments is rare \cite{henin09}.

Derivation of different lipid headgroup conformational ensembles in liquid state has been
inconclusive due to lack of suitable experimental data and tools to interpret the conformational ensembles.

Besides the structure, also ion binding may regulate biophysical activity of especially negatively
charged lipid headgroups \cite{seelig90}. Monovalent cation (except Lithium) binding to zwitterionic PC and anionic PS headgroups is
very weak, while multivalent ion binding is stronger but still weak \cite{cevc90,tocanne90,roux90,catte16,antila19}. The ion binding affinity
data for PE is more scarce \cite{marra85}, but large differences to PC would be surprising.
Negatively charged lipids are suggested to bear same cation binding constants than zwitterionic lipids, but
the amount of bound ions to negatively charged membranes would still be larger because
the concentration of cations in the vicinity of membranes would be higher  \cite{seelig90}.
%is stronger
%Based on the electrometer concept and other data is has been suggested that \cite{seelig90}
%\begin{displayquote}
%  {\it ''(i) Ca$^{2+}$ binds to neutral lipids (phosphatidylcholine, phosphatidylethanolamine) and negatively charged lipids
%    (phosphatidylglycerol) with approximately the same binding constant of K = 10-20 M$^{-1}$; \\
%    (ii) the free Ca$^{2+}$
%    concentration at the membrane interface is distinctly enhanced if the membrane carries a negative surface
%    charge, either due to protein or to lipid; \\
%    (iii) increased inter-facial Ca$^{2+}$ also means increased amounts
%    of bound Ca$^{2+}$ at neutral and charged lipids; \\
%    (iv) the actual binding step can be described by a Langmuir
%    adsorption isotherm with a 1 lipid:1 Ca$^{2+}$ stoichiometry, provided the interfacial concentration C$_M$, is
%    used to describe the chemical binding equilibrium.''}
%\end{displayquote}
On the other hand, anionic PS lipids are proposed chelate with calcium ions \cite{hauser85,feigenson86,roux91}.
In simulations, the cation binding affinity to PC and PS membranes is typically overestimated \cite{catte16,antila19},
which can be improved by applying the ECC to the partial charges of the force fields \cite{melcr18,melcr19}.

Here, we use open collaboration and order parameters of glycerol backbone and headgroup
to evaluate the accuracy of PE and PG heagroup structures, and the cation binding affinity

to anionic membranes containing PG lipids in the current MD simulation force fields.
The force field giving the best description for glycerol backbone and headgroup structures
of PC, PS, PG and PE headgroups (CHARMM36) reproduces the essential differences in order parameters
between these headgroups, and therefore enables the analysis of structural differences between the
headgroups.

%\begin{figure}[]
%  \centering
%  \includegraphics[width=9.0cm]{../Figs/lipids.pdf}
%  \caption{\label{lipids}
%    Chemical structures and labels for the headgroup carbons.
%  }
%\end{figure}

\section{Methods}
\subsection{Experimental C--H bond order parameters}
The headgroup and glycerol backbone C--H bond order parameter magnitudes and signs of POPE and POPG
were determined by measuring the chemical-shift resolved dipolar splittings
with a R-type Proton Detected Local Field (R-PDLF) experiment~\cite{dvinskikh04} and
%The corresponding order parameter signs were measured with a
S-DROSS experiments~\cite{gross97} using natural abundance $^{13}$C solid state NMR spectroscopy
as described previously \cite{ferreira13,ferreira16}.
%The experiments were done in a Bruker Avance III 400 spectrometer operating at a $^1$H Larmor frequency of 400.03 MHz.
%Magic angle spinning (MAS) of the sample was used at a frequency of 5.15 kHz (R-PDLF experiment) and 5 kHz (S-DROSS experiment).
%The following experimental setups were used.
POPE and POPG powder were purchased from Avanti polar lipids.
The NMR experiments were identical to our previous work~\cite{antila19}.
\todo{Is this enough and correct, or should we repeat some methods from the NMRlipidsIVps paper?}
The POPE experiments were recorded at 310~K and POPG experiments at 298~K, where the bilayers are in the liquid disordered phase \cite{marsh13}.

Absolute values of the headgroup and glycerol backbone order parameters from PE and PG lipids are measured
previously using $^2$H NMR~\cite{seelig76,gally81,wohlgemuth80,borle85}. Because also the order parameter
signs bear essential information about the lipid structures \cite{botan15,ollila16}, we measured the
magnitudes and signs of POPE and POPG C--H bond headgroup and glycerol backbone order parameter in liquid phase
using the 2D-RPDLF and S-DROSS experiments, as described previously \cite{ferreira13,ferreira16,antila19}.
For POPE, the glycerol backbone and $\alpha$-carbon peaks in INEPT spectra were assigned based on
previously measured POPC spectra~\cite{ferreira13} and
the $\beta$-carbon peak was assigned based on $^{13}$C chemical shift table for amines available
at \url{https://www.chem.wisc.edu/areas/reich/nmr/c13-data/cdata.htm} (Fig. \ref{POPEspectra}).
For POPG, the glycerol backbone peaks in INEPT spectra were assigned based on
previously measured POPC spectra~\cite{ferreira13}, while $\alpha$ and  $\gamma$-carbon peaks
\todo{How were these assigned?} (Fig. \ref{POPGspectra}). The numerical value of the $\beta$-carbon
order parameter could not be determined, because its peak overlapped with the g$_2$ peak from glycerol backbone in POPG.
However, the order parameter of $\beta$-carbon is expected to be clearly smaller than for g$_2$
based on previous $^2$H NMR measurements \cite{wohlgemuth80,gally81,borle85}.
Therefore, the beginning of the S-DROSS curve gives the sign for g$_2$ order parameter and end for $\beta$ (Fig. \ref{POPGspectra} (E)).
This is confirmed with SIMPSON calculations using negative value for g$_2$ and positive value for $\beta$ order parameter (Fig. \ref{POPGsimpson}).
\todo{Details to be checked by Tiago}.

\subsection{Molecular dynamics simulations}

Molecular dynamics simulation data were collected using
the Open Collaboration method \cite{botan15}, with
the NMR\-lipids Project blog (\url{nmrlipids.blogspot.fi}) and
GitHub repository (\url{github.com/NMRlipids/NMRlipidsIVotherHGs})
as the communication platforms.
The simulated systems of pure PE and PG bilayers without additional ions
are listed in Tables \ref{systemsPE} and \ref{systemsPG},
and lipid mixtures with additional ions in Table~\ref{systemsMIX}.
Further simulation details are given in the SI, and
the simulation data are indexed in a
searchable database available at \url{www.nmrlipids.fi},
and in the NMRlipids/MATCH repository (\url{github.com/NMRlipids/MATCH}).

The C--H bond order parameters were calculated directly
from the carbon and hydrogen positions using the definition
\begin{equation}
S_{\rm CH}=\frac{1}{2}\langle 3\cos^2\theta -1 \rangle,
\end{equation}
where $\theta$ is the angle between the C--H bond and the membrane normal
(taken to align with $z$, with bilayer periodicity in the $xy$-plane).
Angular brackets denote average over all sampled configurations.
The order parameters were first calculated averaging over time separately
for each lipid in the system. The average and
the standard error of the mean were then calculated over different lipids.
Python programs that use the MDAnalysis library \cite{agrawal11,gowers16}
used for all atom simulations is available in Ref. \citenum{MATCHgit}
({\tt scripts/calcOrderParameters.py}). For united atom simulations, the trajectories
with hydrogens having ideal geometry were constructed first using either buildH program~\cite{buildH}
or ({\tt scratch/opAAUA\_prod.py}) in  Ref. \citenum{MATCHgit}, and the order parameters were
then calculated from these trajectories. This approach has been tested against trajectories
with explicit hydrogens and the deviations in order parameters are small \cite{buildH,piggot17}.\\
\todo{BuildH program is now cited with a direct link to the GitHub repo. I think that a release to Zenodo would be nice in the final publication.}\\
\todo{Maybe we should also shortly discuss here about the reasons for slight dependence of order parameter values on the method used to reconstruct hydrogens?}\\
The ion number density profiles were calculated using the {\tt gmx density} tool
of the Gromacs sofware package \cite{gromacsMANUAL}.

\subsection{Analysis of molecular dynamics simulation data}
The big data set of MD simulations was analysed in the NMRlipids databank manner.
Unique naming convention for lipid atoms in each force field was defined using the mapping files
and analysis for all simulations indexed in NMRlipids databank manner were performed
using python codes.

\subsection{Analysis of lipid conformations bound to proteins}
Dihedral angles of all available conformations in the PDB databank were calculated using the
API access to the databank.

\clearpage
\section{Results and Discussion}

%\subsection{Headgroup and glycerol backbone order parameters of POPE and POPG from $^{13}$C NMR}
\subsection{Conformational ensembles of different lipid headgroups in bulk bilayer}

To experimentally characterize the differences in lipid headgroup conformational ensembles
in liquid bilayer phase, we measured the C-H bond order parameters magnitudes and signs
from POPG and POPE bilayers, and compared those to our previously published results
for POPC \cite{ferreira13,ferreira16} and POPS \cite{antila19} in figure \ref{HGorderParameters}.
%Our experimental order parameters, including the signs, from different lipid headgroups
%combined with the literature data are shown in Fig. \ref{HGorderParameters}.
%Also the headgroup order parameters of PC lipids are close PE, althought the latter gives systematically
%slightly more positive values (Fig. \ref{HGorderParameters}).
% These could be explained with slightly larger temperature in PE measurements, except for the $\alpha$-carbon with the positive sign, for which the
% more positive value is farther away from zero.
%While the headgroup order parameters are similar for PC and PE lipids,
%PG and PS lipids exhibit distinct values in comparison with other lipids.
The most distinct order parameters are observed for PS and PG headgroups.
In PS, the $\alpha$-carbon order parameter exhibits significant forking
and the $\beta$-carbon has more negative value than in other lipids.
In PG headgroup,
%the $\alpha$-carbon order parameter is similar to PE and PG, while
the $\beta$-carbon order parameter has positive sign in contrast to all the other lipids.
Notably, this has not been observed in traditional $^2$H NMR experiments,
%because absolute value of $\beta$-carbon order parameter is similar in PG, PE and PC lipids and
where only the absolute value of the order parameters are measured~\cite{wohlgemuth80,gally81,borle85}.
The glycerol backbone order parameters are similar for all the lipids, although they move slightly toward
positive values (closer to zero) in the order PC $<$ PE $<$ PS $<$ PG.

%As discussed previously for PC and PS headgroups \cite{ollila16,antila19}, also 
%the headgroup and glycerol backbone order parameters of PE and PG are essentially independent of 
%acyl chain composition, and therefore manifest mainly headgroup chemistry (Fig. \ref{HGorderParameters}).

\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/lipids.pdf}
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPCPSPEPG.eps}
  \caption{\label{HGorderParameters}
    (top) Chemical structure of different lipids.
    (bottom) Headgroup and glycerol backbone order parameters
    from different experiments in lamellar liquid disordered phase.
    The values and signs for POPE (310~K) and POPG (298~K)
    measured in this work, and for POPS (298~K) \cite{antila19} and POPC (300~K) \cite{ferreira13,ferreira16}
    previously using $^{13}$C NMR. The literature values for
    DOPS with 0.1M of NaCl (303~K) \cite{browning80},
    POPG with 10nM PIPES (298~K) \cite{borle85},
    DPPG with 10mM PIPES and 100mM NaCl (314~K) \cite{wohlgemuth80}, 
    DPPE (341~K) \cite{seelig76},
    E.coliPE and E.coliPG (310~K) \cite{gally81}
    are measured using $^2$H NMR. The signs from $^{13}$C NMR are used also for the literature values.
  }
  \todo{The bottom figure could be clarified as Fig. 2 in the NMRlipids IVps paper.}
\end{figure}

To characterize the differences in lipid headgroup conformational ensembles
leading to the distinct order parameters in PS and PG lipids, we calculate the
the distributions of heavy atom dihedral angles from CHARMM36 simulations.
Among available simulation models, this force field has least problems in
reproducing the experimental lipid headgroup and glycerol backbone order parameters in different lipids,
see Refs. \cite{botan15} and \cite{antila19}, and Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}.
Importantly, it captures the experimentally observed distinct order parameters for PS and PG headgroups (Fig. \ref{??}).

show major differences between lipids only for O$_\alpha$-C$_\alpha$ and  C$_\alpha$-C$_\beta$ bonds
near the end of the headgroup (Fig. \ref{structures}).
These probably explain the slight differences in the
angle between headgroup dipole and membrane normal, which decreases 
in the order PG $>$ PE  $>$ PC  $>$ PS, although difference between PC and PE may an artefact as
the $\beta$-carbon order parameter in PC is poorly reproduced by the CHARMM36 force field.
As in previous NMRlipids project results for PC and PS lipids \cite{botan15,antila19},
none of the MD simulation force fields correctly captures all the
headgroup and glycerol backbone order parameters of PE and PG lipids
(Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG})
that would enable a straightforward interpretation of conformational ensembles.
%show wide variation between different force fields
%and none of the force fields reproduce all values within experimental error bars.
%(Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}),
%as observed previously also for PC and PS lipids \cite{botan15,antila19}.
%The Slipid simulations were able to capture the essential differences between PC and PS lipid
%headgroups \cite{antila19}, but this is not the case for PE and PG lipids.
%Without further discussion about poorly performing force fields, we focus on more detailed analysis
%of
Nevertheless, CHARMM36 force field, which gives the results closest to the experiments for all lipids,
captures the essential differences between
PC, PS, PG and PE headroup order parameters  (Fig.~\ref{structures}) with the exception of $\beta$-carbon order parameter of PC
%give the best overall agreement with experiments for
%the headgroup and glycerol backbone order parameters of all PC, PS, PG and PE lipids in this and our previous
%studies \cite{botan15,antila19}. Even thought many order parameters in CHARMM36 simulations are
%not within the experimental error bars, it 
which is too negative when compared with other lipids or experiments \cite{botan15}.
% As already pointed out in our previous study \cite{antila19},

Dihedrals of both O-P bonds can freely rotate in all lipids as all possible angles
are observed in the distributions. For PS, however, some angles are quite unlikely,
possible explaining the more rigid headgroup structures proposed for PS lipids~\cite{browning80,buldt81}.
The eclipsed (0 or 360 degrees) are not present in any other dihedrals, and anti-eclipsed (120 or 240 degrees)
are not present in C$_\alpha$-C$_\beta$ and g$_1$-g$_2$ bonds.

%Characteristic dihedral conformations in PS headgroup are asymmetric conformations
%preferring gauche 270$^\circ$ conformations in N-C$_\beta$-C$_\alpha$-O$_\alpha$ and
%%dihedral exhibits a more asymmetric angle distribution for PS, having largest probability being
%%in , than for the PC headgroup.
%C$_\beta$-C$_\alpha$-O$_\alpha$-P dihedrals. 
%In PG headgroup, the O$_\beta$-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral (corresponding N-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral in other% lipids)
%is mostly in trans conformation, and C$_\beta$-C$_\alpha$-O$_\alpha$-P  has asymmetric tendency to be in
%gauche 60$^\circ$ conformation. 
%Main difference between PC and PE is the lower probability of trans state in C$_\beta$-C$_\alpha$-O$_\alpha$-P PC dihedral,
%which could be a potential reason for the too negative $\beta$ headgroup order parameter in PC.%

In conclusion, the results suggest that free rotations around O-P bonds decouple the
headgroup and glycerol backbone structure and dynamics in similarly in all lipids,
althought this rotation is slower in PS lipids where crossing certain barriers are slower.
Despite significant differences in dihedral distributions near the headgroup end,
all lipids have dihedral angles within approximately the same ranges.
This suggests that all lipids can quite freely arrange to multiple headgroup conformations
when interaction with proteins, ions or other biomolecules.
%the order parameter experiments suggest that the glycerol backbone conformations in all lipids
%and the headgroup conformations in PC and PE lipids are relatively similar, while 
%PS and PG headgropus exhibit distinct conformational sampling.
%The details of sampled conformation are difficult to deduce from order parameters only,
%but the distinct headgroup order parameters
%of PS lipids are previously related to the more rigid structure of the headgroup~\cite{browning80,buldt81,antila19}.

\begin{figure*}[!h]
  \centering
  \includegraphics[width=18.0cm]{../Figs/PCPEPGPSstructures2.png}
  \includegraphics[width=8.0cm]{../Figs/CHARMMfromLIPIDS.eps}
%  \includegraphics[width=8.0cm]{../Figs/PCPEPGPSdihedrals1.png}
%  \includegraphics[width=8.0cm]{../Figs/PCPEPGPSdihedrals2.png}
  \includegraphics[width=8.0cm]{../Figs/PNangleNORM.eps}
  \includegraphics[width=8.0cm]{../Figs/DIHEDRALS.eps}
  \caption{\label{structures}
    Overlayed snapshots and dihedral angle distributions from CHARMM36 simulations of different lipids
    which give the best agreement with experiments.
  }
  \todo{More detailed discussion of this figure is in https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/9} \\
\end{figure*}

\clearpage
\subsection{Lipid conformational ensembles in lipid bilayers with bound ions}
Phosphatidylcholine headgroup conformational ensemble and orientation are unchanged
upon addition of neutral molecules such as cholesterol or sphingomyelin into a bilayer,
while incorporation of charges affect the headgroup dipole vector tilt which can
be detected by measuring C-H bond order parameters of $\alpha$ and $\beta$ carbons \cite{??}.
Such changes have been observed upon addition of charged lipids, proteins, surfactants,
drugs, or ions \cite{??}, but how charges affect the lipid conformational ensemble remains unknown.
%Headgroup order parameters of PC lipids decrease or increase upon addition of positive
%or negative charge into membrane, respectively, while they remain 
%\cite{??}.
%This has been explained by the tilting of the headgroup dipole due to electrostatic
%interactions with bound charges, and has been used to measure the binding affinity
%of ions to membranes \cite{??}.
%The headgroup order parameters of PC lipids can be used to measure the ion binding
%affinity to lipid bilayers because their magnitude is linearly proportional to the amount of bound charge in bilayer \cite{seelig87,catte16}.
%This molecular electrometer concept can be used also for bilayers containing PC lipids mixed with charged lipids \cite{borle85,macdonald87,roux90,antila19}.
%The headgroup order parameters can be used to evaluate MD simulations against experimental data,
%because they can be directly calculated from MD simulations \cite{catte16}.

\begin{figure*}[bt]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPCvsSURFchange.eps}
  \caption{\label{changesWITHsurf}
    Modulation of PC headgroup order parameters and P-N vector angle upon addition of cationic surfactant.
  }
\end{figure*}

\begin{figure}[]
  \centering
  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MCHARMM36.png}
  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
  \caption{\label{DIHSwithSURF}
    Changes in CHARMM36 dihedrals with increasing amount of cationic surfactant.
  }
\end{figure}

The response of PC headgroup order parameters to the addition of cationic surfactants into a
membrane in CHARMM36 simulations are in very good agreement with the experimental data (Fig. \ref{changesWITHsurf})
suggesting that these simulations can be used to interpret how lipid headgroups respond to the
inclusion of charges into a membrane. Analysis of lipid dihedral bond distributions suggest that
the lipid conformational changes upon addition of charge occur in phosphate-oxygen and g$_3$-phosphate-oxygen bonds.

To resolve the lipid headgroup conformational ensembles in mixed membranes we analyzed
headgroup order parameters from mixture of PC and PE or PG and compared the changes
upon mixing to the experimental data \cite{??} (Figs. \ref{??}). However, the currently
available force fields failed to capture the headgroup conformational ensembles in mixed membranes
as we previously observed also for mixtures with PS lipids \cite{antila19}. The best performing force field for
single componen membrane, CHARMM36, overestimated the influcence of PE to PC conformations and
undersestimated response to anionic PG headgroup. The latter may arise from overbinding of sodium
counterions \cite{antila19}.

Also the changes of headgroup order parameters in POPC:POPG (1:1) and (4:1) mixtures
upon addition of CaCl$_2$
%from simulations and experiments \cite{borle85,macdonald87}
in Figs. \ref{changesWITHCaClPG} and \ref{changesWITHCaClPG1PC4} are in line with
conclusions from our previous studies \cite{catte16,melcr18,antila19,melcr20}:  
%The decrease of POPC headgroup order parameters in mixtures with POPG lipids with
%increasing CaCl$_2$ concentration is overestimated in Slipids and Lipid17
%order parameters decrease in the response to bound cations in qualitative agreement with
%experiments, but
calcium binding affinity to membranes is typically overestimated in simulations, but
CHARMM36 with the NBfix correction underestimates the binding affinity, and that the
%On the other hand,
results can be improved by implicit inclusion of electronic polarizability.
Among the available simulations, there are two models that correctly reproduce the
PG headgroup order parameter changes upon addition of calcium, Lipid17 and Slipids,
even though the binding affinity of calcium is overestimated in these simulations.
The response of PC to calcium binding is correctly reproduced by Lipid17ecc model, as
previously reported for other membranes \cite{melcr18,melcr19}.
%to Amber based lipid models
%improves the ion binding affinity leading to a good agreement with experiments.
%(Figs. \ref{changesWITHCaClPG} and \ref{changesWITHCaClPG1PC4})
%
%indicating too strong binding affinity of calcium into the bilayers as previously
%observed for pure PC lipid bilayers and mixtures with PS lipids \cite{catte16,antila19}.
%
%The decrease of $\alpha$-carbon order parameter of PC lipids in PC:PG mixtures as a function
%of calcium concentration is close to experiments CHARMM36 simulations (Fig. \ref{changesWITHCaClPG}),
%but the decrease of $\beta$-carbon order parameter seems to be overestimated.
%However, the $\beta$-carbon order parameter was not actually measured from these samples,
%but they are calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
%On the other hand, the simulations are not converged.
%
%The calcium binding affinity to lipid bilayers with PC and PS lipids
%was recently improved by applying the electronic continuum correction (ECC) to Amber Lipid14/17 force fields \cite{melcr18,melcr20}.
%In this approach, the electronic polarizability is implicitly included in the classical force fields
%by scaling the charges with constant factors \cite{leontyev11}. Here, we make a ECC-POPG force field
%by applying the scaling factors originally used for POPS also to POPG, i.e., we multiply
%charges and Lennard-Jones $\sigma$s of headgroup, glycerol backbone, and carbonyl regions with 
%$f_q$=0.75 and $f_\sigma$=0.89, respectively \cite{melcr20}.
%ECC-POPG model gives a weaker calcium binding affinity (Fig.\ref{CAdensPG}) and better agreement with the experimental
%PC headgroup order parameter changes (Fig. \ref{changesWITHCaClPG}) for POPC:POPG mixtures than the original Lipid17 model
%\todo{to be finished when we have all the data}, indicating that the ECC improves the simulation predictions of
%calcium binding affinity as previously observed for PC and PS lipids \cite{melcr18,melcr20}.

\begin{figure*}[bt]
  \centering
  \includegraphics[width=18.0cm]{../Figs/CHANGESwithCaClPG1PC1.eps}
  \caption{\label{changesWITHCaClPG}
    Modulation of headgroup order parameters of POPC ({\it left}) and POPG ({\it right}) in POPC:POPG (1:1)
    mixture upon addition of CaCl$_2$ in 298 K temperature from experiments \cite{borle85,macdonald87} and simulations.
    The $\beta$-carbon order parameter of POPC (dashed line on top left) is not directly measured but
    calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
    The changes with respect to the systems without CaCl$_2$ are shown for other data than
    for the $\alpha$-carbon of POPG for which experimental order parameter is not available.
    %(right) The headgroup order parameters of PG from PC:PG mixtures as a function CaCl$_2$
    %concentration from experiments \cite{borle85} and CHARMM36 simulations.
  }
\end{figure*}

\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/CAdensPG1PC1.eps}
  \caption{\label{CAdensPG}
    Calcium ion density profiles along membrane normal from simulations of POPC:POPG (1:1) mixtures with different force fields.
  }
\end{figure}

Changes in lipid conformational ensembles characterized by distributions of heavy atom dihedral angles
are very small upon binding of calcium or addition of cationic surfactants to membranes (Figs. \ref{??}).
Yet, such changes are sufficient to tilt the headgroup dipole angle and reproduce the experimentally observed
order parameter changes in $\alpha$ and $\beta$ carbons.

\begin{figure}[]
  \centering
  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MCHARMM36.png}
  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
  \caption{\label{DIHSwithCA}
    Changes in POPC CHARMM36 dihedrals with increasing amount of CaCl$_2$.
  }
\end{figure}

\begin{figure}[]
  \centering
  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_MCHARMM36.png}
  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
  \caption{\label{DIHSwithCA}
    Changes in POPG CHARMM36 dihedrals with increasing amount of CaCl$_2$.
  }
\end{figure}

%Simulations also qualitatively reproduce the experimentally observed decrease of POPG
%%Experimental data for the
%$\beta$-carbon order parameter
%%of POPG shows a rapid decrease with increasing
%upon addition of CaCl$_2$,
%%concentrations up to 10 mM
%suggesting that the PG headgroup response is qualitatively captured by simulations, in contrast
%to PS in our previous work \cite{antila19}.
%and more modest decrease with larger concentrations (Fig. \ref{changesWITHCaClPG}) \cite{borle85}.
%This behaviour is similar to that of $\beta$-carbon order parameters of POPC,
%but essentially different than observed for POPS, where $\beta$-carbon order parameters increases with
%addition of calcium \cite{antila19}. Experimentally measured changes of PG $\alpha$-carbon order parameters upon addition of
%calcium are not available.
%%upon addition of calcium in experiments,
%%exhibiting 
%%The behaviour is qualitatively different than in the $\beta$-carbon of PS lipids which increase upon addition of CaCl$_2$.

Lipid17 and Slipids force fields correctly capture the PG $\beta$-carbon order parameter response to CaCl$_2$
even thought the binding affinity was too large based on the comparison of PC headgroup order parameter changes with experiments.
%While applying ECC to Lipid17 improved the PC headgroup order parameter response and binding affinity,
The response of PG $\beta$-carbon order parameter to calcium is too weak in CHARMM36 and ECC-lipid17 force fields,
but decrease of the order parameter is associated with decreased P-N vector angle in all
force fields. Even thought the change in headgroup orientation upon addition of calcium is smaller
in PG lipids than in PC, the conformational changes in PG are more significant because also
dihedral angle distribution around O$_\alpha$-C$_\alpha$ bond is affected by calcium binding (Fig. ??).
However, because none of the force fields simulatneously capture both the PG conformational ensemble and calcium binding,
more accurate force fields are required for more detailed studies of the effect of ions to PG conformational enembles.
%and PG
%The response of PG $\alpha$-carbon order parameters to CaCl$_2$ differs between force fields,
%but experimental data to evaluate these predictions is not available. 
%Simulations also qualitatively reproduce the experimentally observed decrease of POPG
%%Experimental data for the
%$\beta$-carbon order parameter
%%of POPG shows a rapid decrease with increasing
%upon addition of CaCl$_2$,
%%concentrations up to 10 mM
%suggesting that the PG headgroup response is qualitatively captured by simulations, in contrast
%to PS in our previous work \cite{antila19}.
%and more modest decrease with larger concentrations (Fig. \ref{changesWITHCaClPG}) \cite{borle85}.
%This behaviour is similar to that of $\beta$-carbon order parameters of POPC,
%but essentially different than observed for POPS, where $\beta$-carbon order parameters increases with
%addition of calcium \cite{antila19}. Experimentally measured changes of PG $\alpha$-carbon order parameters upon addition of
%calcium are not available.
%%upon addition of calcium in experiments,
%%exhibiting 
%%The behaviour is qualitatively different than in the $\beta$-carbon of PS lipids which increase upon addition of CaCl$_2$.

% The result is similar to the $\sim$200~ns simulations with PC lipids in previous work \cite{catte16}.
%However, when simulation was continued for $\mu$s, the binding affinity substantially increased
%and interpretation was that calcium overbinds to PC lipid in CHARMM36. Therefore, the
%conclusion seems to be similar here, although the new NBfix parameters may complicate
%the situation \todo{The status of NBfix parameters in these simulations should be checked.}.

%The $\beta$-carbon order parameter of PG exhibits a rapid decrease with small CaCl$_2$ concentrations
%and a more modest decrease with larger concentrations in experiments \cite{borle85} (Fig. \ref{PSPGchangesWITHCaCl}).
%The rapid decrease with CaCl$_2$ is observed but overestimated in CHARMM36 simulation
%with POPC:POPG 1:1 mixture, but not in 4:1 mixture \todo{This is little bit weird, should be checked.}.

%\todo{We still need more data to finish the discussion. More detailed discussion is in https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12}

\subsection{Protein bound lipid conformations }
\begin{figure*}[!h]
  \centering
  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPCfromPDB.eps}
  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPEfromPDB.eps}
  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPGfromPDB.eps}
  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPSfromPDB.eps}
  \caption{\label{structures}
    Dihedral distributions from simulations and lipid structures in PDB.
  }
\end{figure*}

\clearpage
\section{Conclusions}

% Tables may be be put in the text as floats.
% Here is an example of the general form of a table:
% Fill in the caption in the braces of the \caption{} command. Put the label
% that you will use with \ref{} command in the braces of the \label{} command.
% Insert the column specifiers (l, r, c, d, etc.) in the empty braces of the
% \begin{tabular}{} command.
%
% \begin{table}
% \caption{\label{} }
% \begin{tabular}{}
% \end{tabular}
% \end{table}

% If you have acknowledgments, this puts in the proper section head.
\begin{acknowledgments}
AP is grateful to the Centro de
Supercomputación de Galicia (CESGA) for use of the Finis
Terrae computer
    %     Put your acknowledgments here.
\end{acknowledgments}
%\newpage
%\appendix
%\begin{center}
%{\bf SUPPLEMENTARY INFORMATION}
%\end{center}

%\section{Measurements of order parameter sign}

%Fig. \ref{PShgSIGNS} summarizes the experimental results on the order parameter sign
%measurement for POPS sample. The experimental protocol is the same used in Ref. \citenum{ferreira16}.
%In (a) you see the headgroup region of the INEPT spectrum where alpha and beta are
%identified. In (b) you have the R-PDLF slices for alpha and beta where you see one single
%splitting for beta (which gives an order parameter equal to 0.12), and for alpha a superposition
%of a large splitting (order parameter equal to 0.09) and a very small splitting which cannot be
%calculated. On the bottom you have the S-DROSS slices of these two carbons. The grey lines show a
%random collection of slices from noise such that it gets clear what is significant. The S-DROSS
%slice for beta clearly shows that the order parameter is negative. The slice for alpha shows that
%the higher order parameter is positive and suggests that the smaller order parameter is negative
%(from the deviation towards negative values in the longer t1 times).
%\begin{figure}[]
%  \centering
%  \includegraphics[width=9.0cm]{../Figs/PShgSIGNS.pdf}
%  \caption{\label{PShgSIGNS}
%    Experimental results for sign measurement for POPS sample
%  }
%\end{figure}

%The results updated with SIMPSON simulations for the SDROSS profiles
%are shown in Fig. \ref{PShgSIGNSsimpson}. The value for the smaller
%alpha order parameter is taken from Fig 3 in Ref. \citenum{roux91},
%because resolution in 13C NMR experiments was nor high enough to determine
%numerical value for this. The plots in Fig. \ref{PShgSIGNSsimpson} (c) show
%the following. The error bars and points are the experimental SDROSS data.
%The thick lines are SIMPSON simulations. The simulations were done by using
%the order parameter for beta equal to -0.12 and for alpha one order parameter
%equal to 0.09 and the other equal to -0.02 (black) or 0.02 (grey).
%Since the black lines agree with experimental data, we conlude that
%the order parameters for $\beta$ carbon are -0.12 and for $\alpha$
%order parameters are 0.09 and -0.02.
%\begin{figure}[]
%  \centering
%  \includegraphics[width=9.0cm]{../Figs/PShgSIGNSsimpson.pdf}
%  \caption{\label{PShgSIGNSsimpson}
%    Experimental results for sign measurement for POPS sample
%  }
%\end{figure}

%\section{Dihedrals}
%\begin{figure*}[]
%  \centering
%  \includegraphics[width=8.0cm]{../Figs/dihed1.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed2.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed3.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed4.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed5.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed6.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed7.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed8.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed9.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed10.png}
%  \caption{\label{dihedrals}
%    Experimental results for sign measurement for POPS sample
%  }
%\end{figure*}

% Create the reference section using BibTe
\bibliography{refs.bib}

%\newpage
%\section{APPENDIX: The NMR results reported by Tiago Ferreira}

\listoftodos

\end{document}
%
% ****** End of file aiptemplate.tex ******